---
title: "Data manipulation using dplyr"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

Author: Team BioSakshat

Last update: June 2017

<p>Copyright &copy; 2017 BioSakshat, Inc. All rights reserved.</p>


## Lets begin

### Load dplyr library
```{r}
library(dplyr);
```

### Load Gene Expression Data
```{r}
gen=read.table("_site/data/Day2/GenExp_data.txt", header = TRUE, stringsAsFactors = FALSE);
head(gen);
dim(gen);
```

### Filter rows with filter()
filter() allows you to select a subset of rows in a data frame.

**Filter coding genes.**
```{r}
temp=filter(gen, Genetype=="Coding");
head(temp);
dim(temp);
```

*Q1) Filter only those genes which are in pathway P2. How many such genes are there*

**Filter those genes whose gene expression values on Day1 (D1) > 1.5 (up regulated).**
```{r}
filter(gen, D1> 1.5);
```

**Filter those genes whose gene expression values on Day1 (D1) > 1.5 and Day2 (D2) > 1.**
```{r}
filter(gen, D1> 1.5, D2>1);
```

*Q2) Filter those genes whose expression on Day1 > 1.5 and belong to Pathway P2.*

**Filter genes whose D1 or D2 expression is > 1.5**
```{r}
# Using OR condition: |
temp=filter(gen, D1 > 1.5 | D2 > 1.5);
head(temp);
dim(temp);
```
**Filter genes from pathway P2 whose D1 or D2 expression is > 1.5**
```{r}
filter(gen, D1 > 1.5 | D2 > 1.5, Path=="P2")
```


### Piping using %>%
Piping minimizes the use of variables. It cleans the code. Results from left hand site is piped to right hand side. 

**How many genes are there in Pathway P2**
```{r}
filter(gen, Path=="P2") %>% nrow
```

1. Filter those genes which below to pathway P2.
2. For the filtered genes, filter further whose D1 expression are > 1
```{r}
filter(gen, Path=="P2") %>% filter(., D1 > 1) %>% nrow
# . is optional if used as 1st argument
filter(gen, Path=="P2") %>% filter(D1 > 1) %>% nrow
```

### Arrange rows with arrange()
```{r}
# Arrange genes first as per Pathway followed by expression values on D1
arrange(gen, Path, D1) %>% head
# Arrange genes first as per Pathway followed by expression values on D1 in descending order
arrange(gen, Path, desc(D1)) %>% head
```

### Select rows position wise using slice()
```{r}
# Select 25th to 30th row
slice(gen, 25:30); 
```


### Select column using select()
```{r}
# Select D1, D5 and Path column
select(gen, D1, D5, Path) %>% head;
# Exclude D2 column
select(gen, -D2) %>% head;
# Exclude D2, D5 column
select(gen, -c(D2,D5)) %>% head;
# Column name starts with
select(gen, starts_with("D")) %>% head;
# Column name ends with 3
select(gen, ends_with("3")) %>% head;
# Column name contains "e"
select(gen, contains("e")) %>% head;
# Select all column from D1 to Path
select(gen, D1:Path) %>% head;
# Select 2nd and 4th column
select(gen, c(2,4)) %>% head
```

### Extract unique rows using distinct()
```{r}
# Select unique entries in D1 column
select(gen, D1) %>% distinct %>% nrow
# Select D2 and Path column then find unique rows
select(gen, D2, Path) %>% distinct %>% nrow
```

### Add new columns with mutate() and transmute() function
```{r}
# mutate(gen,difD1D2=D1-D2): Adds a column named as “difD1D2”, which consists of difference between values of D1 and D2 gene expressions, to the existing dataset.
mutate(gen, diffD1D2 = D1-D2) %>% head
# mutate(gen,difD1D2=D1-D2,pdifD1D2=difD1D2*100) Use newly created column difD1D2 to create another pdifD1D2. So, you can create and reuse newly created columns.
mutate(gen, diffD1D2 = D1-D2, pdiffD1D2=diffD1D2*100) %>% head
mutate(gen, diffD1D2 = D1-D2, pdiffD1D2=diffD1D2*100, meanexp=rowMeans(select(gen,D1,D2,D3,D4,D5))) %>% head
# Just keeps the new columns only.
transmute(gen, difD1D2=D1-D2, pdifD1D2 = difD1D2 * 100) %>% head
```

### Summarise values with summarise()
```{r}
summarise(gen, D1mean = mean(D1), D2mean = mean(D2))
```

### Grouped operations
```{r}
# Group genes as per pathway
gengrp=group_by(gen, Path);
# For every group then do some operation
summarise(gengrp, n())
summarise(gengrp, mean(D1), mean(D2), mean(D3))
summarise(gengrp, mD1=mean(D1), mD2=mean(D2), mD3=mean(D3)) %>% filter(.,mD1>0.05) %>% select(.,Path,mD1);
summarise(gengrp, OverExp_Genecount_D1=sum(D1>1.5), OverExp_Genecount_D2=sum(D2>1.5), OverExp_Genecount_D3=sum(D3>1.5))
```

### Joins two table
Currently dplyr supports following join types:

  * inner_join()
  * left_join()
  * right_join()
  * semi_join()
  * anti_join()
  * full_join() 

Create Pathway information data frame which we will use to demonstrate join
```{r}
# Select one gene from each pathway, for demo purpose
tempgen = group_by(gen,Path) %>% slice(1);
# Create Pathway information. Note that we dont have pathway information for P4
Pathwayinfo=data.frame(Path=c("P1","P2","P3","P5"), Name=c("Glycolysis","Gluconeogenesis","Purine metabolism","Apotopsis"), ID=c("ID0752","ID3251","ID2567","ID4568"));
Pathwayinfo;
```

```{r}
tempgen;
Pathwayinfo;

# Return all rows from x where there are matching values in y (i.e. matching x rows)
# Return all columns from x and y. 
# If there are multiple matches between x and y, all combination of the matches are returned.
inner_join(tempgen, Pathwayinfo, by="Path");


# Return all rows from x
# All columns from x and y. 
# Rows in x with no match in y will have NA values in the new columns. If there are multiple matches between x and y, all combinations of the matches are returned.
left_join(tempgen, Pathwayinfo, by="Path");

# Return all rows from y
# all columns from x and y. 
# Rows in y with no match in x will have NA values in the new columns. If there are multiple matches between x and y, all combinations of the matches are returned.
right_join(tempgen, Pathwayinfo, by="Path");


# Return all rows and all columns from both x and y. 
# Where there are not matching values, returns NA for the one missing.
full_join(tempgen, Pathwayinfo, by="Path");


# Return all rows from x where there are matching values in y, keeping just columns from x.
semi_join(tempgen, Pathwayinfo, by="Path");


# Return all rows from x where there are not matching values in y, keeping just columns from x.
anti_join(tempgen, Pathwayinfo, by="Path");
```

### References
1. https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html


