---
title: "ggplot2"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```


## Plot using ggplot2

### Scatter plot
```{r}
library(ggplot2);
library(dplyr);
head(diamonds);

# This will just loads the dataset. No plot will be printed until you add the geom layers.
ggplot(diamonds);
# Here we have added x and y axis. Still no plot of data. 
ggplot(diamonds, aes(x=carat, y=price)); 
```

### Add geom_point() layer
```{r}
# Add geom_point layer. Each layer can be added by +
ggplot(diamonds, aes(x=carat, y=price))+ geom_point() 
```

### Add geom_smooth() layer, linear modeling
```{r}
# 
ggplot(diamonds, aes(x=carat, y=price))+ geom_point() + geom_smooth();
# se=FALSE removes confidence bands
ggplot(diamonds, aes(x=carat, y=price))+ geom_point() + geom_smooth(se=FALSE);
```

### Explore aesthetic parameter "col"
```{r}
# Color the points w.r.t. cut column
ggplot(diamonds, aes(x=carat, y=price, color=cut))+ geom_point() + geom_smooth();
```

### Assign aes() to individual layer
```{r}
ggplot(diamonds, aes(x=carat, y=price))+ geom_point(aes(color=cut)) + geom_smooth(); # We have removed aes to smooth layer
```

### Explore aesthetic parameter "shape"
```{r}
ggplot(diamonds, aes(x=carat, y=price))+ geom_point(aes(color=cut, shape=clarity)) + geom_smooth(); # We have removed aes to smooth layer
```

### Add axis lables and plot title using labs()
```{r}
ggplot(diamonds, aes(x=carat, y=price))+ geom_point(aes(color=cut, shape=clarity)) + geom_smooth() + labs(title="Scatterplot", x="Carat", y="Price");
```

### Change color pelette
```{r}
ggplot(diamonds, aes(x=carat, y=price))+ geom_point(aes(color=cut, shape=clarity)) + geom_smooth() + labs(title="Scatterplot", x="Carat", y="Price")+scale_colour_brewer(palette = "Set1") ;
```

### Save the ggplot object and then print.
```{r}
g=ggplot(diamonds, aes(x=carat, y=price))+ geom_point(aes(color=cut, shape=clarity)) + geom_smooth() + labs(title="Scatterplot", x="Carat", y="Price");
print(g);
```

### The Theme
We can use theme() function to adjust size of labels. Parameters plot.title, axis.text.x, axis.text.y,  axis.title.x,  axis.title.y can be set using element_text() function

```{r}
g <- ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + labs(title="Scatterplot", x="Carat", y="Price");

gg1 <- g + theme(plot.title=element_text(size=30, face="bold"), 
                  axis.text.x=element_text(size=15), 
                  axis.text.y=element_text(size=15),
                  axis.title.x=element_text(size=25),
                  axis.title.y=element_text(size=25));

print(gg1);
```

### Adjusting the legend title
You can change legned title. Based on the type of legend ggplot2 provides different function. For a legend representing color and if the color attribute is derived from discrete values, use scale_color_discrete() function. If legend correspond to shape and discrete use scale_shape_discrete(). Other functions are 
scale_shape_continuous(name="legend title"). For fill  attribute: scale_fill_continuous(name="legend title")
```{r}
gg2 <- gg1 + scale_color_discrete(name="Cut of diamonds") + scale_shape_discrete(name="clarity attribute");
print(gg2);
```

### Facet_wrap
```{r}
# Split based on "cut" column. plot will be distributed in n*3 layouts
gg <- ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + labs(title="Scatterplot", x="Carat", y="Price")
gg3 = gg + facet_wrap( ~ cut, ncol=3);
print(gg3);
```

### Bar charts
```{r}
ggplot(diamonds,aes(x=cut))+geom_bar();
ggplot(diamonds,aes(x=cut))+geom_bar(aes(fill=clarity));
ggplot(diamonds,aes(x=cut))+geom_bar(aes(fill=clarity)) + coord_flip();
# Suppose for every cut class, plot mean depth. x-axis: cut class while y-axis: mean(depth)
tempdf=group_by(diamonds, cut) %>% summarise(meandepth=mean(depth))
ggplot(tempdf, aes(x=cut, y=meandepth))+geom_col();
# geom bar for continious data
ggplot(diamonds, aes(x=carat))+geom_histogram()
```

### Density plot
```{r}
ggplot(diamonds, aes(carat))+ geom_density();
ggplot(diamonds, aes(carat, colour=cut))+ geom_density();
ggplot(diamonds, aes(carat, colour=cut, fill= cut))+ geom_density();
ggplot(diamonds, aes(carat, colour=cut, fill= cut))+ geom_density(alpha=0.1);
# Stacked density plots:
ggplot(diamonds, aes(carat, colour=cut, fill= cut))+ geom_density(alpha=0.1, position="stack");
```

### Box plot
```{r}
ggplot(diamonds, aes(x=cut, y=table)) + geom_boxplot();
ggplot(diamonds, aes(x=cut, y=table)) + geom_boxplot(notch = TRUE);
ggplot(diamonds, aes(x=cut, y=table, colour=cut)) + geom_boxplot(notch = TRUE);
ggplot(diamonds, aes(x=cut, y=table)) + geom_boxplot(fill = "white", colour = "#3366FF"); 
ggplot(diamonds, aes(x=cut, y=table)) + geom_boxplot(outlier.colour = "red", outlier.shape = 1);
```

### References
1. http://r-statistics.co/ggplot2-Tutorial-With-R.html
2. http://ggplot2.tidyverse.org/reference/

