---
title: "Conditional Statements & Loops"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

Author: Team BioSakshat

Last update: June 2017

<p>Copyright &copy; 2017 BioSakshat, Inc. All rights reserved.</p>

## Conditional execution
Syntax:  
if (expr_1)  
{  
    expr_2    
    ......  
    ......  
}else{    
  expr_3    
  ......  
  ......  
}  

```{r}
x=sample(1:50,10);
sum(x);
if(sum(x)>500)
{
  print("Sum of x is greater than 500");
}else{
  print("Sum of x is less than 500");
}
```

## Repetitive execution using for loop
```{r}
for(i in 5:10)
{
  print(i);
}
```

```{r}
for(i in c(10,15,17,16,25))
{
  print(i);
}
```

```{r}
x=c(10,15,17,16,25);
for(i in x)
{
  print(i);
}
```

```{r}
x=c(10,15,17,16,25);
for(i in 1:length(x))
{
  print(c(i,x[i]));
}
```

Store squares of values of above vector in separate vector s using for loop
```{r}
s = NULL;
for(i in 1:length(x))
{
  s[i] = x[i] ^ 2;
}
x;
s;
```
Note: Usually we don’t need to use for loop to perform operations on a single vector. The vector s created using above for loop can simply be created using command, s<- x^2. But for loop is very useful to perform operations on
multiple columns and rows of a matrix or data frame.

### Task
Create a matrix of 5 rows for values 1 to 50. Calculate the means of all rows in a matrix and store them in a vector m.
```{r}
mat=matrix(sample(1:50, 50, replace = TRUE), nrow=10);
mat;
m=NULL;
for(i in 1:nrow(mat))
{
  m[i] = mean(mat[i,]);
}
m;
```





