---
title: "Matrix and Array"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

Author: Team BioSakshat

Last update: June 2017

<p>Copyright &copy; 2017 BioSakshat, Inc. All rights reserved.</p>

## Matrix

### Create a matrix from vector using dim attributes
```{r}
x=21:70;
dim(x)=c(10,5); # 10 row, 5 col
x;
```


### Create matrix using matrix()
```{r}
x=21:70;
matrix(x, ncol=10); # 10 col
m=matrix(x, ncol=10, byrow = TRUE); # Fill elements row wise
```

### Functions that can be applied to a matrix
```{r}
# Data structure of m
str(m); 
# Dimension of m
dim(m);
# Number of rows
nrow(m);
# Number of columns
ncol(m);
# Column means
colMeans(m);
# Row means
rowMeans(m);
# Transpose
t(m);
# Fetch column names
colnames(m);
# Fetch row names
rownames(m);
# head
head(m);
```

### Assign column/row names
```{r}
colnames(m)=paste("C",1:ncol(m),sep="");
rownames(m)=paste("R",1:nrow(m),sep="");
m;
```

### Fetch elements from a matrix using indices
```{r}
m;
# 1st row, 3rd column
m[1,3];
# 1st tp 3rd row and 3rd to 5th column
m[c(1:3),c(3:5)];
# 1st row
m[1,];
# 3rd column
m[,3];
# 3rd element. Here matrix is converted to vector and 3rd element is returned
m[3];
```

### Fetch element using index matrix
```{r}
m;
index=c(2,3,3,4,4,5);
m[index]
```

### Fetch elements from a matrix using row and column names
```{r}
m;
m["R1","C3"];
```

### Insert rows and columns by rbind() and cbind()
```{r}
m1=matrix(1:10, nrow=5);
m1;
m1=rbind(m1, 999);
m1;
m1=rbind(m1, c(2,3));
m1;
m1=cbind(m1, 1:5);
m1;
m1=cbind(m1, 10:20);
m1;
```

### Create matrix using rbind() and cbind()
```{r}
x=seq(2, 4, length=10);
y=1:10;
x;
y;
cbind(x,y);
rbind(x,y);
```

### Delete row(s) and column(s)
```{r}
m;
m=m[-1,];
m;
m=m[,-3];
m;
```

### Check for matrix
```{r}
is.matrix(m);
```

### Conditional statements on matrix
```{r}
m;
m%%5==0;
which(m%%5==0, arr.ind = TRUE);
```

**Task: Now Go to [Task page](task.html) and finish Matrix**
