---
title: "Solution"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

Author: Team BioSakshat

Last update: June 2017

<p>Copyright &copy; 2017 BioSakshat, Inc. All rights reserved.</p>

# Solutions

## Vector creation

**1. Create a vector x13 with values 2, 3, 4, 5, 6**

x13=c(2,3,4,5,6)

**2. Create a vector x14 with values 2.0, 2.1, 2.2, 2.3, 2.4, .., 4**

x14=seq(from=2,to=4, by=0.1)

**3. Create a vector x15 with 10 random values between 4 and 6**

x15=sample(4:6,10, replace=T)

**4. Create a vector x16 with repeated values 3, 4, 5, 3, 4, 5, 3, 4, 5**

x16=rep(c(3,4,5),times=3)

**5. create x17 with repeated values 7,7,7,8,8,8,9,9,9**

x17=rep(c(7,8,9),each=3)

**6. Create a vector x18 with 10 random values between 20 and 30**

x18=sample(20:30,10)

**7. Create a vector x19 with 10 normally distributed random values**

x19=rnorm(10)

**8. Create a vector x20 with values of vectors x13 and x16 followed by 3, 5,10**

x20=c(x13,x16,3,5,10)

## Fetching vector elements

**9. Create a vector x21 with values 33,55,66,88,99 and fetch its 3rd, 5th and 2nd values**

x21=c(33,55,66,88,99)

x21[c(3,5,2)]

**10. Fetch its values from 1 to 4**

x21[1:4]

**11. Fetch values of x21 vector excluding 2nd and 3rd elements**

x22=x21[-c(2,3)]

x22

**12 Fetch last element of x21 using length()**

x21[length(x21)]

## Vector manipulation
**13. Create a vector x23 with values 5, 7, 6, 8, 1, 4.**

x23=c(5,7,6,8,1,4)

**Delete 1st and last elements.  **

x23=x23[-c(1,length(x23))]

**Reset the value of second element to 12.**

x23[2]=12

**Add value 0 at the beginning of a vector x23**

x23=c(0,x23)
 
## Vector arithmetic

**14. Calculate Variance**

Write the arithmetic expression to calculate variance of a vector. Cross check your result using var() function. 

Formula: Variance= sum((x-mean(x))^2) /n-1 where n is total number of elements.

x=c(5,7,6,8,1,4)

xvar=sum((x-mean(x))^2)/(length(x)-1)

xvar

var(x)

**15. Distance between atoms**

Given x, y, z coordinates of two atoms. Atom1 (1.2, 2.3, 3.4) and Atom2 (4.5, 5.6, 6.7). Find distance between 2 atoms. Formula: sqrt((x2-x1)2+(y2-y1)2+(z2-z1)2)

atom1=c(1.2,2.3,3.4)

atom2=c(4.5,5.6,6.7)

dist=sum((atom1-atom2)^2)

dist

**16. Find out the numbers between 1 to 100, which are divisible by 2 or 3.**

x=1:100

x[which(x%%2==0 & x%%3==0)]

## Matrix

**17 Create matrix **

Create a matrix from a vector cosisting of numbers from 1 to 12 with 3 columns

mt=matrix(1:12,ncol=3)

**18. Fetch 2nd row. Fetch 3rd column**

mt[2,]

mt[,3]

**19. Fetch the value 6**

mt[2,2]

**20. Fetch the value 8 and 12**

mt[4,c(2,3)]

**21. Fetch the value 7, 8, 11 and 12**

mt[c(3,4),c(2,3)]

## Data Frame

**22. Create a data frame of gene expression data such that**

First column,"Genes", is a character vector of 6 gene names (G1, G2, ..., G6). 

Second column, "C1" is a numeric vector of 6 random values from 3 to 5. 

Hint: Generate random numbers using function sample(). Use R help to see the syntax of sample.

Third column, "C2" is a numeric vector of 6 random values from 3 to 5. 

Fourth column, "T1" is a numeric vector of 6 random values from 5 to 7. 

Fifth column, "T2" is a numeric vector of 6 random values from 5 to 7. 

Sixth column, "Pathway" is a character vector of which first 3 represent one pathway "P1" and other 3 represent pathway "P2". 



genes=paste("G",1:6,sep="")

C1=sample(3:5,6, replace=T)

C2=sample(3:5,6, replace=T)

T1=sample(5:7,6, replace=T)

T2=sample(5:7,6, replace=T)

pathway=rep(c("P1","P2"),each=3)

exp=data.frame("Genes"=genes,"C1"=C1,"C2"=C2,"T1"=T1,"T2"=T2,"Pathway"=pathway)

**23. Fetch values of column D2**

exp$T2

**24. Fetch values of gene G2**

exp[2,]

**25. Fetch value of gene G3 from C2**

exp[3,3]

**26. Delete column C2**

exp$C2=NULL

**27. Insert column C3, which should be numeric vector of 6 random values from 3 to 5. **

exp$C3=sample(3:5,6, replace=T)

**28. Find mean of column C1**

mean(exp$C1)

## Tasks on iris data set

**1. Open iris data set file using read.table() and store in a variable name "iris_data"**

iris_data=read.table("iris.txt",header=T)

**2. Check the structure of "iris_data". Note the column names. How many categories are there in column named as "Species". Note the names of species.**

str(iris_data)

levels(iris_data$Species)

**3. How many rows and columns are there?**

dim(iris_data)

**4. How many observations (rows) are there for each species?  **

**a. Number of rows with Species as setosa**

**b. Number of rows with Species as virginica**

**c. Number of rows with Species as versicolor** 

table(iris_data$Species)

**5. Find the mean of all sepal lengths.**

mean(iris_data$Sepal.Length)

**6. Find the mean of sepal lengths in Species setosa.**

library(dplyr)

setosa=filter(iris_data, Species=="setosa")

mean(setosa$Sepal.Length)

**7. What is the overall correlation between Sepal length and Petal length?**

cor(iris_data\$Sepal.Length, iris_data\$Petal.Length)

**8. What is the correlation between Sepal width and Petal width of Species virginica?**

virginica=filter(iris_data, Species=="virginica")

cor(virginica\$Sepal.Width, virginica\$Petal.Width)

**9. Find the difference between sepal lengths of species setosa and versicolor. What is the mean difference between them?**

versicolor=filter(iris_data, Species=="versicolor")

sldiff= setosa\$Sepal.Length - versicolor\$Sepal.Length

mean(sldiff)

**10. Carry out the t-test between sepal lengths of species setosa and versicolor. Is it statistically significant? **

Assumptions:
Sepal lengths of two species are normally distributed
Their variances may not be equal
Observations are not paired
Two sided t-test, we are just interested to know if sepal lenghts between species are significantly different or not.


t.test(setosa\$Sepal.Length, versicolor\$Sepal.Length)

Based on p-value we reject null hypothesis at 95 % confidence interval
Hence sepal lengths of species setosa and versicolor are significantly different

Note: Please see the help for t.test to find out various options for t-test


## Visualization
Plot1: Scatter plot of Sepal length vs Petal length of all 150 flowers, color according to species/plants.
```{r, eval=FALSE}
library(gdata)
irisd=read.xls("iris.xls");
```

```{r, eval=FALSE}
plot(irisd$SL,irisd$PL,col=rep(c("red","blue","green"),each=50),pch=15,xlab="Sepal Length",ylab="Petal Length",cex.lab=1.2,main="Sepal Vs Petal Length",cex.axis=1.2)
legend(4.2,7,c("Setosa","Versicolor","virginia"),col=c("red","blue","green"),pch=15)
```
Plot2: Barplot showing distribution  of Sepal lengths among 6 classes of flowers (3 plants and 2 colors). 
```{r, eval=FALSE}
barplot(t(table(irisd$SP,irisd$Col)),beside=T,col=c("red","blue"),ylim=c(0,50),cex.axis=1.5,cex.name=1.5)
legend(1,50,c("Red","Blue"),pch=15,col=c("red","blue"))
```
Plot3: Multi panel plot showing the histogram of SL, PL, SW, PW of all 150 flowers. 
```{r, eval=FALSE}
par(mfrow=c(2,2))
hist(irisd[,1],xlab="Sepal Length",ylab="Frequency",main="",cex.lab=1.2,col="gray")
hist(irisd[,2],xlab="Sepal width",ylab="Frequency",main="",cex.lab=1.2,col="gray")
hist(irisd[,3],xlab="Petal Length",ylab="Frequency",main="",cex.lab=1.2,col="gray")
hist(irisd[,4],xlab="Petal Width",ylab="Frequency",main="",cex.lab=1.2,col="gray")
par(mfrow=c(1,1))
```

Plot4: Box plot showing SL, SW, PL, PW distribution along with a line joining their mean lengths.
```{r, eval=FALSE}
boxplot(irisd[,1:4],col=c("red","green","blue","orange"),main="Length Distribution",cex.axis=1.2,ylab="Length (cm)",cex.lab=1.2)
l=summarise(irisd,mean(SL),mean(SW),mean(PL),mean(PW))
lines(c(1,2,3,4),l,lwd=4,col="brown")
legend(3.5,8,c("Mean\nLength"),lty=1,lwd=4,col="brown")
```

Plot5: Probability density plot of Sepal lengths among three different categories of plants.

```{r, eval=FALSE}
# First filter
library(dplyr)
set_SL=filter(irisd,SP=="setosa") %>% select(SL)
versi_SL=filter(irisd,SP=="versicolor") %>% select(SL)
virgi_SL=filter(irisd,SP=="virginica") %>% select(SL)

# Convert them to vector
set_SL=set_SL$SL
versi_SL=versi_SL$SL;
virgi_SL=virgi_SL$SL;


# use density to plot
plot(density(set_SL),xlim=c(4,10),lwd=2,xlab="Length (cm)",main="Prob. Density of Sepal Length across different flowers")
lines(density(versi_SL),col="red",lwd=2)
lines(density(virgi_SL),col="green",lwd=2)
legend(8,1.2,lwd=2,c("Setosa","Versicolor","Virginica"),col=c("black","red","green"))
```

Plot6: 3D plot showing distinct clustering of flowers in terms of SL, SW and PL. Different colors for different plants.

```{r, eval=FALSE}
library(rgl)
plot3d(irisd[,1:3],col=rep(c("red","green","blue"),each=50))
```

Plot7: Scatter plot matrix showing a global view of the distribution of SL, SW, PL and PW across 3 plants.

```{r, eval=FALSE}
pairs(irisd[,1:4],col=rep(c("red","green","blue"),each=50))
```


Plot8: Heatmap showing clustering of flowers in terms of their SL, SW, PL and PW properties.

```{r, eval=FALSE}
m=as.matrix(irisd[,1:4]);
library("gplots")
heatmap.2(m, trace="none")
```

Plot9: Pie chart showing the number of flowers in 6 categories (3 plants and 2 colors)

```{r, eval=FALSE}
#Subset species by colors
library(dplyr)

Set_B=filter(irisd, SP=="setosa" & Col=="Blue" )
Set_R=filter(irisd, SP=="setosa" & Col=="Red" )
Ver_B=filter(irisd, SP=="versicolor" & Col=="Blue" )
Ver_R=filter(irisd, SP=="versicolor" & Col=="Red" )
Vir_B=filter(irisd, SP=="virginica" & Col=="Blue" )
Vir_R=filter(irisd, SP=="virginica" & Col=="Red" )

#Create vector of counts in each subset
irisp=c(nrow(Set_B),nrow(Set_R),nrow(Ver_B),nrow(Ver_R),nrow(Vir_B),nrow(Vir_R))

pie(irisp, labels=irisp, col=c("Red", "Cyan", "yellow", "Blue", "Green", "Magenta"))
legend(0.9,1,c("Set_B","Ver_B","Vir_B","Set_R","Ver_R","Vir_R"),pch=15, col=c("Red","Yellow","Green", "Cyan","Blue","Magenta"))
```

Plot10: Dot chart showing clear distribution of SL among 3 plants. 
```{r, eval=FALSE}
dotchart(irisd$SL,group=irisd$SP, main="Sepal Length across 3 Plants")
```


